**Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces**

### 2020-03-02 submission/manuscript.Rmd
Prist part of discussion.

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

@@

# Calculating the mean and standard deviation for each environment and season
S.obs_season_f <- filter(estimators_indices_metadata, station=="F") %>%
  group_by(season) %>%
  summarise(mean=mean(S.obs), sd=sd(S.obs))
S.obs_season_fcym <- filter(estimators_indices_metadata, station=="FCyM") %>%
  group_by(season) %>%
  summarise(mean=mean(S.obs), sd=sd(S.obs))
S.obs_season_fcam <- filter(estimators_indices_metadata, station=="FCaM") %>%
  group_by(season) %>%
  summarise(mean=mean(S.obs), sd=sd(S.obs))
S.obs_season_fca <- filter(estimators_indices_metadata, station=="FCa") %>%
  group_by(season) %>%
  summarise(mean=mean(S.obs), sd=sd(S.obs))
```

@@

The microbial taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}). Three phyla were characteristic for *Archaea*: *Crenarchaeota*, *Nanoarchaeota* and *Thermoplasmatota*. Their contribution  was larger only in seawater communities during December 2017 (\autoref{community}).  Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}). Generally, at higher taxonomic ranks (phylum-class) epiphytic and seawater microbial communities were composed of similar bacterial taxa. Seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*. Communities associated with *Cymodocea nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018. In addition, communities from invaded and noninvaded *Caulerpa cylindracea* were similar and characterized by same groups as seawater and *Cymodocea nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}). Larger differences between environments and host species could be observed at lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).

@@

# Remove Chloroplast and Mitochondria sequences and subtract their number from higher taxonomic levels to which
# they belong
chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
community <- mutate_at(community, 5:ncol(community), list(~case_when(
  rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
  rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
  rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
  TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Chloroplast")) %>%
  mutate_at(5:ncol(.), list(~case_when(
    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
    TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Mitochondria")) %>%
  select(-ATCC_1, -ATCC_5, -NC_1) %>%
  mutate(`23`=`23_1`+`23_2`) %>%
  select(-`23_1`, -`23_2`) %>%
  group_by(taxlevel) %>%
  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
  ungroup()

@@

*Cyanobacteria* related sequences were comprising, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}). Higher proportions were found for *Cymodocea nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *Caulerpa cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}). Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed. Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of and sequences without known relatives within *Cyanobacteriales* and *Cyanobacteriia* (\autoref{cyano}). In addition, seasonal changes in surface associated communities were observed with *Cyanobacteriales* and *Cyanobacteriia* no relative sequences comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* and in spring and summer (\autoref{cyano}).

@@

Sequences classified as *Bacteroidota* were comprising, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}). Similarly to *Cyanobacteria*, large differences in the taxonomic composition between seawater and surface associated communities were found (\autoref{bactero}). The seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*. In contrast, surface associated communities were characterized by *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae*. Surface associated *Bacteroidota* communities were very diverse as could be observed in the the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).

@@

On average, *Alphaproteobacteria* were in comparison to other high rank taxa the largest taxonomic group, comprising `r alpha_all_whole$mean` ± `r alpha_all_whole$sd` \si{\percent} of all sequences (\autoref{alpha}). In accordance to previous taxa, high differences between seawater and surface associated communities were observed. Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria* and no relative  *Hyphomonadaceae*. Representatives of no relative *Rhodobacteraceae* were comprising in on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}).

@@

Sequences related to *Gammaproteobacteria* were comprising, on average, `r gamma_all_whole$mean` ± `r gamma_all_whole$sd` \si{\percent} of all sequences (\autoref{gamma}). Similarly to previous taxa, large taxonomic differences between seawater and surface associated communities were found. Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicocus*. Beside these two groups specific to all three epiphytic communities, *Cymodocea nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *Caulerpa cylindracea*. In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).

@@

# Calculating the mean and standard deviation for contrasting season of Caulerpa cylindracea
desulfo_station_season_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  filter(station=="FCaM" | station=="FCa") %>%
  mutate(season=str_replace(season, "Winter", "Summer")) %>%
  mutate(season=str_replace(season, "Spring", "Autumn")) %>%
  group_by(station, season) %>%
  summarise(mean=mean(abundance), sd=sd(abundance))
```
In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *Caulerpa cylindracea*. On average they were comprising `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences. While seawater and *Cymodocea nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *Caulerpa cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}). In addition, *Caulerpa cylindracea* associated communities were characterized by higer proportions in Winter and Summer (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$sd` \si{\percent}) in comparison to Autumn and Spring (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$sd` \si{\percent}). The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).

## Discussion

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = T)
opts_chunk$set("message" = FALSE)

@@

# Calculating the mean and standard deviation for each environment and season
S.obs_season_f <- filter(estimators_indices_metadata, station=="F") %>%
  group_by(season) %>%
  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
S.obs_season_fcym <- filter(estimators_indices_metadata, station=="FCyM") %>%
  group_by(season) %>%
  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
S.obs_season_fcam <- filter(estimators_indices_metadata, station=="FCaM") %>%
  group_by(season) %>%
  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
S.obs_season_fca <- filter(estimators_indices_metadata, station=="FCa") %>%
  group_by(season) %>%
  summarise(mean_S.obs=mean(S.obs), sd_S.obs=sd(S.obs),
            mean_S.chao1=mean(S.chao1), sd_S.chao1=sd(S.chao1),
            mean_S.ACE=mean(S.ACE), sd_S.ACE=sd(S.ACE),
            mean_eshannon=mean(eshannon), sd_eshannon=sd(eshannon),
            mean_invsimpson=mean(invsimpson), sd_invsimpson=sd(invsimpson))
```

@@

The taxonomic composition of both, macrophyte associated and seawater communities, was dominated by *Bacteria* (`r filter(archaea_bacteria, taxon=="Bacteria")$mean` ± `r filter(archaea_bacteria, taxon=="Bacteria")$sd` \si{\percent}) over *Archaea* (`r filter(archaea_bacteria, taxon=="Archaea")$mean` ± `r filter(archaea_bacteria, taxon=="Archaea")$sd` \si{\percent}) (\autoref{community}). Higher relative abundances of chloroplast related sequences were only observed in surface associated communities, with higher values in Autumn/Winter (`r filter(chloroplast, season=="Winter")$mean` ± `r filter(chloroplast, season=="Winter")$sd` \si{\percent}) in comparison to Spring/Summer (`r filter(chloroplast, season=="Summer")$mean` ± `r filter(chloroplast, season=="Summer")$sd` \si{\percent}) (\autoref{supp-chloroplast}). Generally, at higher taxonomic ranks (phylum-class) epiphytic and seawater microbial communities were composed of similar bacterial taxa. Seawater communities were mainly comprised of *Actinobacteriota*, *Bacteroidota*, *Cyanobacteria*, *Alphaproteobacteria*, *Gammaproteobacteria* and *Verrucomicrobiota*. Communities associated with *Cymodocea nodosa* were consisted of same groups with the addition of *Planctomycetota* whose contribution was higher in summer 2018. In addition, communities from invaded and noninvaded *Caulerpa cylindracea* were similar and characterized by same groups as seawater and *Cymodocea nodosa* communities with the addition of *Desulfobacterota* (\autoref{community}). Larger differences between environments and host species could be observed at lower taxonomic ranks (\autoref{cyano} -- \ref{desulfo}).

@@

# Remove chloroplast and mitochondrial sequences and subtract their number from higher taxonomic levels to which
# they belong
chloroplast <- filter(community, str_detect(taxon, "^Chloroplast$"))$rankID
mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
community <- mutate_at(community, 5:ncol(community), list(~case_when(
  rankID==str_extract(chloroplast, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Chloroplast"],
  rankID==str_extract(chloroplast, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Chloroplast"],
  rankID==str_extract(chloroplast, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Chloroplast"],
  TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Chloroplast")) %>%
  mutate_at(5:ncol(.), list(~case_when(
    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
    TRUE ~ .))) %>%
  filter(!str_detect(taxon, "^Mitochondria")) %>%
  select(-ATCC_1, -ATCC_5, -NC_1) %>%
  mutate(`23`=`23_1`+`23_2`) %>%
  select(-`23_1`, -`23_2`) %>%
  group_by(taxlevel) %>%
  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
  ungroup()

@@

*Cyanobacteria* related sequences were comprising, on average, `r cyano_all_whole$mean` ± `r cyano_all_whole$sd` \si{\percent} of total sequences (\autoref{cyano}). Higher proportions were found for *Cymodocea nodosa* (`r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCyM" & season=="Autumn")$sd` \si{\percent}) and *Caulerpa cylindracea* (invaded [ (`r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCaM" & season=="Autumn")$sd` \si{\percent})] and noninvaded [ (`r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$mean` ± `r filter(cyano_station_season_whole, station=="FCa" & season=="Autumn")$sd` \si{\percent})]) associated communities in autumn and for seawater communities in winter (`r filter(cyano_station_season_whole, station=="F" & season=="Winter")$mean` ± `r filter(cyano_station_season_whole, station=="F" & season=="Winter")$sd` \si{\percent}). Large taxonomic differences between surface associated and seawater cyanobacterial communities were observed. Seawater communities were mainly comprised of *Cyanobium* and *Synechococcus*, while surface associated communities were consisted of *Pleurocapsa* and sequences without known relatives within *Cyanobacteriia* (\autoref{cyano}). In addition, seasonal changes in surface associated communities were observed with *Pleurocapsa* and no relative *Cyanobacteriia* comprising larger proportions in autumn and winter and *Acrophormium*, *Phormidesmis* and no relative *Nodosilineaceae* in spring and summer (\autoref{cyano}).

@@

Sequences classified as *Bacteroidota* were comprising, on average, `r bactero_all_whole$mean` ± `r bactero_all_whole$sd` \si{\percent} of all sequences (\autoref{bactero}). Similarly to *Cyanobacteria*, large differences in the taxonomic composition between seawater and surface associated communities were found (\autoref{bactero}). The seawater community was characterized by the NS4 and NS5 marine groups, uncultured *Cryomorphaceae*, uncultured *Flavobacteriaceae*, NS11-12 marine group, *Balneola*, uncultured *Balneolaceae* and *Formosa*. In contrast, in surface associated communities *Lewinella*, *Portibacter*, *Rubidimonas*, no relative *Saprospiraceae*, uncultured *Saprospiraceae*, no relative *Flavobacteriaceae* and uncultured *Rhodothermaceae* were found. Some groups showed slight seasonal changes such as no relative *Flavobacteriaceae* that were more pronounced from November 2017 until June 2018. In contrast, uncultured *Rhodothermaceae* showed higher proportions from June 2018 until the end of the study period. Surface associated *Bacteroidota* communities were very diverse as could be observed in the the high proportion of taxa that grouped as other *Bacteroidota* (\autoref{bactero}).

@@

On average, *Alphaproteobacteria* were in comparison to other high rank taxa the largest taxonomic group, comprising `r alpha_all_whole$mean` ± `r alpha_all_whole$sd` \si{\percent} of all sequences (\autoref{alpha}). In accordance to previous taxa, high differences between seawater and surface associated communities were observed. Picoplankton communities were composed mainly of the SAR11 clade, AEGEAN-169 marine group, SAR116 clade, no relative *Rhodobacteraceae*, HIMB11 and OCS116 clade, while surface associated communities were composed in high proportion of no relative *Rhodobacteraceae* and to a lesser degree of  *Pseudoahrensia*, no relative *Alphaproteobacteria*, no relative  *Hyphomonadaceae* and *Amylibacter*. Representatives of no relative *Rhodobacteraceae* were comprising on average `r rhodo$mean` ± `r rhodo$sd` \si{\percent} of all alphaproteobacterial sequences from the epiphytic community (\autoref{alpha}). In addition, *Amylibacter* was detected mainly in *Cymodocea nodosa* from November 2017 until March 2018.

@@

Sequences related to *Gammaproteobacteria* were comprising, on average, `r gamma_all_whole$mean` ± `r gamma_all_whole$sd` \si{\percent} of all sequences (\autoref{gamma}). Similarly to previous taxa, large taxonomic differences between seawater and surface associated communities were found. Seawater communities were mainly comprised of the OM60 (NOR5) clade, *Litoricola*, *Acinetobacter* and the SAR86 clade, while epiphytic communities were mainly composed of no relative *Gammaproteobacteria* and *Granulosicocus*. Beside these two groups specific to all three epiphytic communities, *Cymodocea nodosa* was characterized by *Arenicella*, no relative *Burkholderiales* and *Methylotenera*, while *Thioploca*, no relative *Cellvibrionaceae* and *Reinekea* were more specific to both invaded and noninvaded *Caulerpa cylindracea*. In addition, *Arenicella* was more pronounced in November and December 2017, while no relative *Burkholderiales* and *Methylotenera* were more characteristic for the period form March until May 2018. For the *Caulerpa cylindracea* specific taxa no relative *Cellvibrionaceae* and *Reinekea* showed some seasonality and were characterisitic for samples originating from June to October 2018. In addition, similarly to *Bacteroidota*, a large proportion of the surface associated community was grouped as other *Gammaproteobacteria* indicating high diversity within this group (\autoref{gamma}).

@@

# Calculating the mean and standard deviation for contrasting season of Caulerpa cylindracea
desulfo_station_season_whole <- inner_join(metadata, whole, by=c("ID"="Group")) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  filter(station=="FCaM" | station=="FCa") %>%
  mutate(season=str_replace(season, "Winter", "Summer")) %>%
  mutate(season=str_replace(season, "Spring", "Autumn")) %>%
  group_by(station, season) %>%
  summarise(mean=mean(abundance), sd=sd(abundance))
```

In contrast to previously described high rank taxa, *Desulfobacterota* were specific to *Caulerpa cylindracea*. On average they were comprising `r desulfo_all_whole$mean` ± `r desulfo_all_whole$sd` \si{\percent} of all sequences. While seawater and *Cymodocea nodosa* communities were consisted of only `r filter(desulfo_station_whole, station=="F")$mean` ± `r filter(desulfo_station_whole, station=="F")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCyM")$mean` ± `r filter(desulfo_station_whole, station=="FCyM")$sd` \si{\percent} *Desulfobacterota* sequences, respectively, in the invaded and noninvaded *Caulerpa cylindracea* communities their proportion was `r filter(desulfo_station_whole, station=="FCaM")$mean` ± `r filter(desulfo_station_whole, station=="FCaM")$sd` \si{\percent} and `r filter(desulfo_station_whole, station=="FCa")$mean` ± `r filter(desulfo_station_whole, station=="FCa")$sd` \si{\percent}, respectively (\autoref{desulfo}). In addition, *Caulerpa cylindracea* associated communities were characterized by higer proportions in Winter and Summer (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Summer")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Summer")$sd` \si{\percent}) in comparison to Autumn and Spring (invaded, `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCaM", season=="Autumn")$sd` \si{\percent}; noninvaded, `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$mean` ± `r filter(desulfo_station_season_whole, station=="FCa", season=="Autumn")$sd` \si{\percent}). The community was mainly consisted of no relative *Desulfobacteraceae*, *Desulfatitalea*, no relative *Desulfobulbaceae*, *Desulfobulbus*, no relative *Desulfocapsaceae*, *Desulfopila*, *Desulforhopalus*, *Desulfotalea*, SEEP-SRB4 and uncultured *Desulfocapsaceae* (\autoref{desulfo}).

## Discussion
Surfaces of marine macrophytes are harboring biofilms consisted of diverse microbial taxa [@Egan2013; @Tarquinio2019]. No standard protocol has been developed to study these macophyte associated microbes [@Ugarelli2019]. Different procedures of microbial cells removal from host surfaces were described, such as host tissue shaking [@Noges2010], scraping [@Uku2007] and ultrasonication [@Cai2014]. All these methods showed different removal efficiencies and none was enabling a complete removal of attached microbial cells. In the present study, we applied an earlier developed removal protocol [@Korlevic_in_press, submitted], based on a previous idea of direct cellular lysis [@Burke2009], to ensure an almost complete cell detachment. The application of a direct lysis procedure coupled with a high frequency sampling protocol and Illumina high resolution amplicon sequencing has enabled us to make a detailed description of bacterial and archaeal communities associated with the surfaces of two marine macrophytes, *Cymodocea nodosa* and *Caulerpa cylindracea*.

Observed highest richness values for *Caulerpa cylindracea* (invaded and noninvaded), middle for *Cymodocea nodosa* and lowest for seawater derived communities. Higher values for surface associated communities in comparison to seawater were described earlier for seagrasses [@Ugarelli2019] and could be attributed to a larger set of inhabitable microniches existing on macrophyte surfaces. In addition, highest values observed for *Caulerpa cylindracea* are probably a consequence of sediment derived OTUs that were present only in *Caulerpa cylindracea* communities. *Caulerpa cylindracea* stolon is attached to surface sediment with rhizoids, so the stolon and rhizoids are in a direct contact with the sediment surface. Part of the surface attached *Caulerpa cylnidracea* community is therefore comprised of sediment derived cells that could cause the observed increase in richness. In addition, seasonal richness differences observed for surface attached communities showed slightly higher values in spring and summer could be explained by a higher macrophyte growth in these seasons [M. Najdek, personal communication; @Zavodnik1998; @Ruitton2005]. During active periods macrophytes exhibit a more dynamic chemical interaction with the surface community causing an opening of a larger set of microniches.
